---
title: "Midterm"
author: "Vincent Bianchi"
output: html_document
editor_options: 
  chunk_output_type: console
---

There are five questions (50 total points). Please type in your answers directly in the R Markdown file. After completion, **successfully** knitr it as an html file. Submit <span style="color:red">**both**</span> the html file and the R Markdown file via Canvas. Please name the R Markdown file in the following format: LastName_FirstName_Midterm.Rmd, e.g. Zhao_Zifeng_Midterm.Rmd.


## Boston Housing Dataset
The Boston Housing dataset contains information on 506 census housing tracts in the Boston area. The data is stored in the file `Boston.csv`. It contains 7 variables, `crim`, `chas`, `nox`, `room`, `dis`, `lstat` and `value`. The data description is as follows.

+ `crim`: Per capita crime rate
+ `chas`: Charles River (=Yes if tract bounds river; =No otherwise)
+ `nox`: Nitric oxide concentration (parts per 10 million)
+ `room`: Average number of rooms per house
+ `dis`: Weighted distances to five Boston employment centers
+ `lstat`: Percentage lower status of the population
+ `value`: Median house price in the tract ($10,000s)

Our goal is to build several statistical models such that we can predict the median house price in new tracts based on information such as crime rate, pollution, and number of rooms.

```{r}
rm(list=ls())
```


###  **Q1 [14 points]** Data Visualization and Partition
**Q1(a) [2 points]**
Let's correctly read in the data in `Boston.csv` and name it as `total_data`. 
```{r Q1(a)}
## Write code solution to Q1(a) here

total_data <- read.csv('Boston.csv')

View(total_data)
```

**Q1(b) [4 points]**
Use an **appropriate** plot to investigate the relationship between `value` and `chas`. Does a tract that bounds Charles river have higher median house price?
```{r Q1(b)}
## Write code solution to Q1(b) here

boxplot(total_data$value~total_data$chas)

```

Answer:
Yes, a tract that bounds Charles river does have a higher median house price.


**Q1(c) [4 points]**
Use a histogram to investigate if the dependent variable `value` is right-skewed. If so, further conduct a log-transformation and create a new variable named `log_value`. Make sure `log_value` is created WITHIN the data frame `total_data`.
```{r Q1(c)}
## Write code solution to Q1(c) here

# Histogram of Value 
hist(total_data$value, main='Histogram of Value', xlab='')

# Log Transformation
total_data$log_value  <- log(total_data$value+1)

# Histogram of Log Value 
hist(total_data$log_value, main='Histogram of Log Value', xlab='')


```


**Q1(d) [4 points]**
Let's partition the data in `total_data` into training **(80%)** and test data **(20%)** and store them as `R` objects `train_data` and `test_data` respectively. Use random seed **`set.seed(7)`**!
```{r Q1(d)}
## Write code solution to Q1(d) here
set.seed(7)
total_obs <- dim(total_data)[1]

# Data Partition / Sample Splitting
train_data_indices <- sample(1:total_obs, 0.8*total_obs)
train_data <- total_data[train_data_indices,]
test_data <- total_data[-train_data_indices,]

train_obs <- dim(train_data)[1]
test_obs <- dim(test_data)[2]

```


### **Q2 [6 points]** Linear Regression
**Q2(a) [4 points]**
Fit a linear regression model of the **log scale** `value` w.r.t. all 6 predictors `crim`, `chas`, `nox`, `room`, `dis`, `lstat` using the **training data**, name it `lm_full`.
```{r Q2(a)}
## Write code solution to Q2(a) here

lm_full <- lm(log_value~crim+chas+nox+room+dis+lstat, data=train_data)


```

**Q2(b) [2 points]**
Look at the summary of `lm_full`. What is the $R^2$ of `lm_full`? Which predictors are statistically significant?
```{r Q2(b)}
## Write code solution to Q2(b) here

summary(lm_full)

summary(lm_full)$r.squared


```

Answer:
The $R^2$ of `lm_full` is `r summary(lm_full)$r.squared`. All 6 predictors, `crim`, `chas`, `nox`, `room`, `dis`, and `lstat` are statistically significant. 


### **Q3 [6 points]** GAM
Fit a GAM of the **log scale** `value` w.r.t. all 6 predictors `crim`, `chas`, `nox`, `room`, `dis`, `lstat` using the **training data**, name it `gam_full`. Let's use splines with **df=4** for all **5** numerical predictors, which include `crim`, `nox`, `room`, `dis` and `lstat`.
```{r Q3}
library(gam)

gam_full <- gam(log_value~s(crim,df=4)+s(nox,df=4)+s(room,df=4)+s(dis,df=4)+s(lstat,df=4)+chas,
                data=train_data)

plot(gam_full, col='skyblue')

```


### **Q4 [14 points]** Neural Networks
Fit an NN of **log scale** `value` w.r.t. all 6 predictors `crim`, `chas`, `nox`, `room`, `dis`, `lstat` using the **training data**, name it `nn_full`. For the architecture of NN, let's use one hidden layer with 4 hidden units.

**Q4(a) [4 points]**
Let's generate the **training dataset** that are needed for the estimation of NN using the function `model.matrix()` and store it in `x_train_nn`. In addition, use the `scale()` function to standardize the predictors by centering with mean and scaling with sd.
```{r Q4(a)}
## Write code solution to Q4(a) here

library(neuralnet)

x_train_nn <- model.matrix(~crim+nox+room+dis+lstat+chas, data=train_data)[,-1]


# standardization
x_mean <- apply(x_train_nn, 2, mean)
x_sd <- apply(x_train_nn, 2, sd)
x_train_nn <- scale(x_train_nn, center=x_mean, scale=x_sd)


```

**Q4(b) [2 points]**
Let's further combine the **log scale** `value` with the standardized predictors `x_train_nn` generated in Q4(a). Don't forget to rename the first column of the data frame as `log_value`!
```{r Q4(b)}
## Write code solution to Q4(b) here

# Combining with dependent variable, log_value
x_train_nn <- cbind.data.frame(train_data$log_value, x_train_nn)
colnames(x_train_nn)[1] <- 'log_value'


```

**Q4(c) [2 points]**
Let's generate the **test dataset** that are needed for the out-of-sample prediction evaluation of NN using the function `model.matrix` and store it in `x_test_nn`. Use the `scale()` function to standardize the predictors by centering with mean and scaling with sd as in Q4(a).
```{r Q4(c)}
## Write code solution to Q4(c) here

x_test_nn <- model.matrix(~crim+nox+room+dis+lstat+chas, data=test_data)[,-1]

x_test_nn <- scale(x_test_nn, center=x_mean, scale=x_sd)

# Combining with dependent variable, log_value
x_test_nn <- cbind.data.frame(test_data$log_value, x_test_nn)
colnames(x_test_nn)[1] <- 'log_value'

```

**Q4(d) [6 points]**
Let's fit an NN that has one hidden layers with 4 hidden units and name it `nn_full`. Make sure to use random seed **`set.seed(7)`**!
```{r Q4(d)}
## Write code solution to Q4(d) here
library(neuralnet)
set.seed(7)

nn_full <- neuralnet(log_value~., data=x_train_nn, hidden = 4 )

plot(nn_full)

```


### **Q5 [10 points]** Model Evaluation (Prediction)
**Q5(a) [4 points]**
Use `lm_full`, `gam_full` and `nn_full` to generate predictions for `value` on the test data and store the prediction in `lm_pred`, `gam_pred` and `nn_pred` respectively. Note that for all predictions, make sure to transform the **log-scale** prediction back to the **original scale**.
```{r Q5(a)}
## Write code solution to Q5(a) here

# prediction of lm_full
lm_pred <- exp(predict(lm_full, newdata = test_data))-1

# prediction of gam_full
gam_pred <- exp(predict(gam_full, newdata = test_data))-1

# prediction of nn_pred
nn_pred <- exp(predict(nn_full, newdata = x_test_nn)[,1])-1


```

**Q5(b) [2 points]**
What are the predicted median house values of the **first** tract in the test data by `lm_full`, `gam_full` and `nn_full`, respectively?
```{r Q5(b)}
## Write code solution to Q5(b) here

lm_pred[1]
gam_pred[1]
nn_pred[1]

```

Answer:
The predicted median house values of the first tract in the test data by `lm_full`, `gam_full` and `nn_full` are `r lm_pred[1]`, `r gam_pred[1]`, and `r nn_pred[1]`, respectively.

**Q5(c) [2 points]**
Use the `R` package `forecast` to evaluate the prediction performance of `lm_full`, `gam_full` and `nn_full` on the test data. What are the MAE for `lm_full`, `gam_full` and `nn_full`?
```{r Q5(c)}
library(forecast)

# lm, gam, nn error
lm_err <- accuracy(lm_pred, test_data$value)
gam_err <- accuracy(gam_pred, test_data$value)
nn_err <- accuracy(nn_pred, test_data$value)

lm_err[3]
gam_err[3]
nn_err[3]

```

Answer:
The MAE for `lm_full`, `gam_full` and `nn_full` are `r lm_err[3]` `r gam_err[3]` and `r nn_err[3]`, respectively.


**Q5(d) [2 points]**
Which statistical model do you prefer, `lm_full` or `gam_full` or `nn_full`? Give reasons. 

Answer: 
I prefer the `nn_full` model because it has the lowest Mean Absolute Error. I also like it because it allows for non-linearity and interaction within the model.

