---
title: "Homework 1: Linear Regression"
author: "Vincent Bianchi"
output: html_document
editor_options: 
  chunk_output_type: console
---

There are six questions (30 total points) in this assignment. The minimum increment is 1 point. Please type in your answers directly in the R Markdown file. After completion, **successfully** knitr it as an html file. Submit <span style="color:red">**both**</span>  the html file and the R Markdown file via Canvas. Please name the R Markdown file in the following format: LastName_FirstName_HW1.Rmd, e.g. Zhao_Zifeng_HW1.Rmd.


## Used Car Dataset [9 points]
The used car dataset is the one we analyzed in class. Let's read in the data stored in `UsedCar.csv`.
```{r chunk1}
total_data <- read.csv("./UsedCar.csv", header=T, stringsAsFactors=T)

View(total_data)
```

### **Q1 [2 points]** 
Since the dependent variable `Price` is severely right-skewed, create a log-transformation of `Price` and store it as a new variable named `Log_price` within the data.frame `total_data`.
```{r Q1}
## Write code solution to Q1 here
total_data$Log_price  <- log(total_data$Price+1)

View(total_data)
```


### **Q2 [7 points]** 
Fit a linear regression model of Log_price w.r.t. two predictors Age and Mileage, name it `lm_short`.
```{r Q2}
## Fit the lm_short model here
lm_short <- lm(Log_price~Age+Mileage, data=total_data)

summary(lm_short)
```

**Q2(a) [2 points]** What is the $R^2$ of `lm_short`? What are the (in-sample) MAE and RMSE of `lm_short` at the *original scale*?
```{r Q2(a)}
## Write code solution to Q2(a) here

summary(lm_short)

library(forecast)

err <- accuracy(exp(lm_short$fitted.values)-1, total_data$Price)

err
#MAE
err[3]
#RMSE
err[2]

#R^2 manually
TSS <- sum((total_data$Log_price-mean(total_data$Log_price))^2)
RSS <- sum((total_data$Log_price-lm_short$fitted.values)^2)
1-RSS/TSS

#R^2 extracted
summary(lm_short)$r.squared

```

Answer: 

The $R^2$ of `lm_short` is `r summary(lm_short)$r.squared`
MAE = `r err[3]`
RSME = `r err[2]`

**Q2(b) [2 point]** What is the estimated coefficient of `lm_short` for Age and Mileage?

```{r Q2(b)}
## Write code solution to Q2(b) here
summary(lm_short)

lm_short$coefficients

lm_short$coefficients[2]
lm_short$coefficients[3]

# Age - -1.217e-02
# Mileage -1.757e-06
```

The estimated coefficients of lm_short for Age and Mileage are `r lm_short$coefficients[2]` and `r lm_short$coefficients[3]` , respectively 

**Q2(c) [3 points]** How should we interpret the estimated coefficient of Age at the log-scale and the original scale of Price?

```{r}
round(abs(lm_short$coefficients[2]*100),3)
round(abs(lm_short$coefficients[2]),3)

```

Answer:

# For every one unit increase in a car's age (for every year the car ages) the nominal price (original scale) of the car decreases by `r round(abs(lm_short$coefficients[2]*100),3)`% and would reduce the log scale price r `round(abs(lm_short$coefficients[2]),3)`.

## Car Seat Sales Dataset [21 points]
The car seat sales dataset contains sales of child car seats at 400 different stores and the data is stored in `Carseats.csv`. It contains 9 variables, `Sales`, `CompPrice`, `Income`, `Advertising`, `Population`, `Price`, `ShelveLoc`, `Age` and `Urban`. We would like to build a linear regression model to predict `Sales` at a planned new store. The data description is as follows.

+ `Sales`: Unit sales (in thousands) at each location
+ `CompPrice`: Price charged by competitor at each location
+ `Income`: Community income level (in thousands of dollars)
+ `Advertising`: Local advertising budget for company at each location (in thousands of dollars)
+ `Population`: Population size in region (in thousands)
+ `Price`: Price company charges for car seats at each site
+ `ShelveLoc`: A factor with levels Bad, Good and Medium indicating the quality of the shelving location for the car seats at each site
+ `Age`: Average age of the local population
+ `Urban`: A factor with levels No and Yes to indicate whether the store is in an urban or rural location

###  **Q4 [2 points]**
Which variable is the dependent variable? Which predictors are categorical variables?

Answer:

The Sales variable is the dependent variable. The ShelveLoc and the Urban predictors are categorical variables.

### **Q5 [9 points]**
Let's read in the data and perform visualization to get a better sense of the data.

**Q5(a) [2 points]** Correctly read in the data stored at `Carseats.csv`.

```{r Q5(a)}
## Write code solution to Q5(a) here

Car_seats_data <- read.csv("./Carseats.csv", header=T, stringsAsFactors=T)

Car_seats_data

```

**Q5(b) [2 points]** Produce a scatterplot between `Sales` and `Price`. What is the general pattern from the scatterplot?
```{r Q5(b)}

## Write code solution to Q5(b) here
plot(Car_seats_data$Price, Car_seats_data$Sales)

```

Answer:

# The general pattern from the scatter plot shows a negative linear correlation between Price and Sales. This means, all else equal, we can generally expect on average less total unit sales the more expensive the price of car seat.

**Q5(c) [2 points]** Produce a boxplot between `Sales` and `ShelveLoc`. What is the general pattern from the boxplot?
```{r Q5(c)}

## Write code solution to Q5(c) here

boxplot(Car_seats_data$Sales~Car_seats_data$ShelveLoc)


```

Answer:

# The general pattern we see from the boxplot between Sales and ShelveLoc is that with better quality shelving locations, we expect more overall sales of car seats.


**Q5(d) [3 points]** Produce a histogram of `Sales`. Is `Sales` severely right-skewed?
```{r Q5(d)}
## Write code solution to Q5(d) here

hist(Car_seats_data$Sales, main='Histogram of Sales', xlab='')

# the data looks normally distributed with no obvious severe skewness

```

Answer:

Based on the box plot Sales is not severely right-skewed.

### **Q6 [10 points]**

**Q6(a) [4 points]** Fit a linear regression model of the original scale `Sales` w.r.t. all the predictors available in the dataset, name it `lm_full`. Take a look at the summary of `lm_full`.
```{r Q6(a)}
## Write code solution to Q6(a) here

lm_full <- lm(Sales~CompPrice+Income+Advertising+Population+ Price+ShelveLoc+Age+Urban, data = Car_seats_data)

summary(lm_full)

```

**Q6(b) [2 points]** What is the estimated coefficient of `lm_full` for CompPrice and Price? How should we interpret the estimated coefficients?
```{r Q6(b)}
## Write code solution to Q6(b) here
coeff <- lm_full$coefficients

coeff[2]
coeff[6]

```

Answer:

# The estimated coefficients for CompPrice and Price are `r coeff[2]`and `r coeff[6]`, respectively. We can interpet from this data that for every unit increase in CompPrice, we expect on average for the unit Sales to increase by a factor of 0.0926753. Likewise we can interpret that for every unit increase in Price we can expect on average for the unit Sales to decrease by a factor of -0.0954335. 

**Q6(c) [2 points]** Which predictor(s) are not statistically significant in the model?

Answer: 
Population and Urban are not statistically significant predictors in the model.

**Q6(d) [2 points]** What is the $R^2$ of `lm_full`? What are the (in-sample) MAE and RMSE of `lm_full`?

```{r Q6(d)}

## Write code solution to Q6(d) here

library(forecast)

# RMSE, MAE
error_values <- accuracy(lm_full$fitted.values, Car_seats_data$Sales)

#RMSE
error_values[2]
#MAE
error_values[3]

summary(lm_full)$r.squared

```

Answer: 

$R^2$ of `lm_full` is `r summary(lm_full)$r.squared`
MAE = `r error_values[3]`
RMSE = `r error_values[2]`

